# Homework 1: Data Exploration and Visualization

This Chapter will focus on the topics surrounding Visualizing Data Models

Topics includes:

- Partial effect plots
- Partial dependent plots
- Individual conditional expectation plots
- Accumulated local effect plots
- (Possible) Permutation feature importance
  

# Module 1 homework
1. Create your course portfolio using the instructions.

2. In the HW1.qmd, create another partial effect plot of predictions with x1 ranging from -3 to 3 and x2 equal -0.5 and x3 set to 0. Repeat for all five models.

## Partial Effect Plots
```{r}
install.packages("pak")
pak::pak("thomasgstewart/ds6021b")
require(ds6021b)

expanded_grid_for_df_mod_01 <- expand.grid(x1 = seq(-3, 3, 0.5), x2 = -0.5, x3 = 0)

# First prediction model
plot(expanded_grid_for_df_mod_01$x1, f1(expanded_grid_for_df_mod_01, x1 = expanded_grid_for_df_mod_01$x1, x2 = -0.5, x3 = 0), title = "First Prediction Model", xlab = "x1", ylab = "Black Box Model 1=F1 for f1(expanded_grid_for_df_mod_01)", type = "p")

# Second prediction model
plot(expanded_grid_for_df_mod_01$x1, f2(expanded_grid_for_df_mod_01, x1 = expanded_grid_for_df_mod_01$x1, x2 = -0.5, x3 = 0), title = "Second Prediction Model", xlab = "x1", ylab = "Black Box Model 2=F2 for f2(expanded_grid_for_df_mod_01)", type = "b")

# Third prediction model
plot(expanded_grid_for_df_mod_01$x1, f3(expanded_grid_for_df_mod_01, x1 = expanded_grid_for_df_mod_01$x1, x2 = -0.5, x3 = 0), title = "Third Prediction Model", xlab = "x1", ylab = "Black Box Model 3=F3 for f3(expanded_grid_for_df_mod_01)", type = "b")

# Fourth prediction model
plot(expanded_grid_for_df_mod_01$x1, f4(expanded_grid_for_df_mod_01, x1 = expanded_grid_for_df_mod_01$x1, x2 = -0.5, x3 = 0), title = "Fourth Prediction Model", xlab = "x1", ylab = "Black Box Model 4=F4 for f4(expanded_grid_for_df_mod_01)", type = "o")

# Fifth prediction model
plot(expanded_grid_for_df_mod_01$x1, f5(expanded_grid_for_df_mod_01, x1 = expanded_grid_for_df_mod_01$x1, x2 = -0.5, x3 = 0), title = "Fifth Prediction Model", xlab = "x1", ylab = "Black Box Model 5=F5 for f5(expanded_grid_for_df_mod_01)", type = "s")```

3. Create another plot that compares predictions for the following using all five models:

 * x1 ranging from -3 to 3 and x2 equal -0.5 and x3 set to 0
 * x1 ranging from -3 to 3 and x2 equal -0.5 and x3 set to 1

 ```{r}
 # Generate predictions for ALL models

# x3 = 0
y1_0 <- f1(x1 = seq(-3, 3, 0.5), x2 = -0.5, x3 = 0)
y2_0 <- f2(x1 = seq(-3, 3, 0.5), x2 = -0.5, x3 = 0)
y3_0 <- f3(x1 = seq(-3, 3, 0.5), x2 = -0.5, x3 = 0)
y4_0 <- f4(x1 = seq(-3, 3, 0.5), x2 = -0.5, x3 = 0)
y5_0 <- f5(x1 = seq(-3, 3, 0.5), x2 = -0.5, x3 = 0)

# x3 = 1
y1_1 <- f1(x1 = seq(-3, 3, 0.5), x2 = -0.5, x3 = 1)
y2_1 <- f2(x1 = seq(-3, 3, 0.5), x2 = -0.5, x3 = 1)
y3_1 <- f3(x1 = seq(-3, 3, 0.5), x2 = -0.5, x3 = 1)
y4_1 <- f4(x1 = seq(-3, 3, 0.5), x2 = -0.5, x3 = 1)
y5_1 <- f5(x1 = seq(-3, 3, 0.5), x2 = -0.5, x3 = 1)

# Create side-by-side plotting area

par(mfrow = c(1,2))

plot(
  seq(-3, 3, 0.5),
  y1_0,
  type = "l",
  col = "black",
  ylim = range(c(y1_0, y2_0, y3_0, y4_0, y5_0)),
  main = "All Models Combined (x3 = 0)",
  xlab = "x1",
  ylab = "Predictions for Black Box Models (1-5)=(F1-F5)"
)

# lines for all Black models f1-f5 for x3 = 0 and x3 = 1
lines(seq(-3, 3, 0.5), y2_0, col = "red", lwd = 2, lty = 2)
lines(seq(-3, 3, 0.5), y3_0, col = "blue", lwd = 2, lty = 3)
lines(seq(-3, 3, 0.5), y4_0, col = "green", lwd = 2, lty = 4)
lines(seq(-3, 3, 0.5), y5_0, col = "purple", lwd = 2, lty = 5)

legend(
  "topright",
  legend = c("Model 1 = F1", "Model 2 = F2", "Model 3 = F3", "Model 4 = F4", "Model 5 = F5"),
  col = c("black", "red", "blue", "green", "purple"),
  lty = c(1,2,3,4,5)
)

plot(
  seq(-3, 3, 0.5),
  y1_1,
  type = "l",
  col = "black",
  ylim = range(c(y1_1, y2_1, y3_1, y4_1, y5_1)),
  main = "All Models Combined (x3 = 1)",
  xlab = "x1",
  ylab = "Predictions for Black Box Models (1-5)=(F1-F5)"
)

lines(seq(-3, 3, 0.5), y2_1, col = "red", lty = 2)
lines(seq(-3, 3, 0.5), y3_1, col = "blue", lwd = 2, lty = 3)
lines(seq(-3, 3, 0.5), y4_1, col = "green", lwd = 2, lty = 4)
lines(seq(-3, 3, 0.5), y5_1, col = "purple", lwd = 2, lty = 5)

legend(
  "topright",
  legend = c("Model 1 = F1", "Model 2 = F2", "Model 3 = F3", "Model 4 = F4", "Model 5 = F5"),
  col = c("black", "red", "blue", "green", "purple"),
  lty = c(1,2,3,4,5)
)
 ```

4. What do we learn from the figures in question 3?

From the figures in question 3, it becomes easy to learn that the model predictions vary significantly across the five different models (F1 to F5) when x1 ranges from -3 to 3, with x2 held constant at -0.5 and x3 set to either 0 or 1. These predictions varies uniquely for each black model value from F1-F5 and x3 values changing and with these variations we can gain an understanding how each model creates different outcomes and graph/model behaviors regardless of utilizing the same input variable x1 from -3 to 3 with incrementation at the value of 0.5. Altogether, this figure helps us understand that there will always be differences in predictions and we need a model by model comparison to better understand these predictions and what is does to the interpretations and decisions we make based on these models.